\chapter{GL-117 Installation}
\label{chap:installation}

This chapter describes the requirements of \emph{GL-117} and its installation
concerning esp. the libraries required to compile and execute the game.

\section{Requirements}
\label{sec:requirements}

\emph{GL-117} requires Linux/Unix or MSWindows as operating system as well as
properly installed versions of the following libraries:
\begin{itemize}
\item{OpenGL or MesaGL: graphics library, 3D engine}
\item{GLU or MesaGLU: utilities for GL}
\item{GLUT or MesaGLUT: a toolkit that provides keyboard and mouse support}
\item{SDL (optional): the Direct Media Layer library has similar features of GLUT plus joystick
support and basic sound processing}
\item{SDL\_mixer (optional): a library that provides advanced multichannel sound
support and music}
\end{itemize}

Installation of \texttt{SDL} and \texttt{SDL\_mixer} is optional, however strongly recommended.

\section{Downloading GL-117}
\label{sec:downloading_gl117}

The latest \emph{GL-117} release is currently available for download at
\texttt{http://www.heptargon.de} in \texttt{.tar.gz} or \texttt{.zip} format.
Try to get a suitable binary package for your system, so you needn't recompile the game.
Some Linux distributions (Debian, PLD) already include the latest game releases and provide an update
mechanism to download and install the game via internet.
Using MSWindows there is already a prebuilt executable. 

Now unpack the archive file with your favourite tool.
For MSWindows I can recommend \texttt{7-zip}, if you are searching for free software that can handle multiple formats.

To extract the files on command line please look at the file suffix:
\begin{verbatim}
unzip gl-117-x.y.z-src.zip
tar zxvf gl-117-x.y.z-src.tar.gz
tar jxvf gl-117-x.y.z-src.tar.bz2
\end{verbatim}
with \texttt{x.y.z} meaning the \emph{GL-117} version
number. For last minute updates and release-specific building and
install instructions, make sure to have a look at the
\texttt{README} and \texttt{INSTALL} files.


\section{Linux/Unix installation}
\label{sec:linux_installation}

If you got a binary package, you will only
need the libraries \texttt{GL, GLU, GLUT} and \texttt{SDL, SDL\_mixer} as
described above and install them, e.g. for \texttt{rpm} files enter:
\begin{verbatim}
 rpm -i SDL-*.rpm
 rpm -i gl-117-*.rpm
\end{verbatim}

In order to \textit{compile} \emph{GL-117} you will also have to install the development
versions \texttt{-devel} of the libraries above (except \texttt{SDL\_mixer}).
To compile \emph{GL-117} (as \textit{root}) do the following steps in the \texttt{gl-117-x.y.z-src}
directory:
\begin{verbatim}
 ./configure
 make
\end{verbatim}

The \texttt{configure} script will check for the required libraries and will output
a \texttt{Makefile}, which can be invoked using the \texttt{make} command.
After compiling \emph{GL-117} successfully, you will find a binary called \texttt{gl-117}
in the \texttt{src} directory.\\
To really \textit{install} \emph{GL-117} please use:
\begin{verbatim}
 make install
\end{verbatim}
This will copy the binary to your binary directory (e.g. \texttt{/usr/local/bin})
and the rest of data files to your data directory (e.g. \texttt{/usr/local/share}).
Any files that require output permissions will be stored in the user's home directory,
exactly \texttt{\$HOME/.gl-117}.\\
This step will require write permissions in the binary and data directories.
However, without \textit{root} permissions you may customize these directories using for example
\begin{verbatim}
 ./configure --prefix='/home/tom/gl-117'
 make
 make install
\end{verbatim}

If you encounter any problems while building or executing the game, please read the \texttt{FAQ}
file or write a mail to \texttt{tom.drexl@gmx.de}.

\section{MSWindows installation}
\label{sec:windows_installation}

First, you might have to install \texttt{GL, GLU, GLUT} and \texttt{SDL, SDL\_mixer}.
Look into your system directory, that is generally
\begin{verbatim}
 \WINDOWS\SYSTEM    for Windows9x/ME
 \WINDOWS\SYSTEM32  for WindowsNT/2000/XP
\end{verbatim}
You will need the files \texttt{opengl32.dll, glu32.dll, glut32.dll, sdl.dll, sdl\_mixer.dll}
there.
If one is missing, please search the internet and move the file there.
That's it. Execute the binary \texttt{gl-117.exe} stored in a directory with the same name.

If you had already an earlier version of \emph{GL-117} you might want to use your
old pilots with the new version of the game. Therefore copy the old
\texttt{saves} directory to the new version.

This game is *NOT* meant to be a Windows game! Do not complain if there is no proper
installer available! Do not complain, if you are not able to unpack and install the game!
There are plenty other and better commercial action games.

\section{Running GL-117}
\label{sec:running_gl117}

At startup, \emph{GL-117} tries to read the files \texttt{conf} and \texttt{conf.interface}
from the user's home directory (Linux/Unix) or the \texttt{saves} directory (MSWindows).
If there are none, the game will try out some screen settings and store the files.
\begin{verbatim}
 Linux/Unix      $HOME/.gl-117/conf
                 $HOME/.gl-117/conf.interface
 MSWindows       GL-117-INSTALLDIR/saves/conf
                 GL-117-INSTALLDIR/saves/conf.interface
\end{verbatim}

You may edit these file using your favourite text editor and adjust the settings
to your system. Since version \textit{1.1}, \emph{GL-117} provides a full menu based configuration
of these files.

If you lack a hardware accelerated video card, please turn down the
quality to 0 or 1. Further acceleration can be achieved negligating fullscreen mode
and choosing a lower resolution.\\
Just delete the files \texttt{conf} and/or \texttt{conf.interface}
if you want to reset to the initial settings.

Typical performance values:

 AMD K2 500 AGP1x TNT, quality=0, view=50: FPS$>$30\\
 Athlon 1400 GeForce2, quality=3, view=70: FPS$>$30\\
 P4 2.4GHz GeForce4, quality=4, view=110: FPS$>$30
